{\bf Practical Concurrency Testing} \\ \normalsize \vspace{1em} \begin{tabular}{c} {\em or, How I Learned to Stop Worrying and Love the Exponential Explosion} \end{tabular}

### 2018-07-19 thesis/tm.tex
fix conflict abort discussion to account for causality

Up to three abort codes are considered depending on the testing options listed in the previous section.
All abort codes are simultaneously added to both lists,
unless already present on {\tt xabort\_codes\_ever},
in which case not added to {\tt xabort\_codes\_todo} to avoid duplicate work.
\begin{itemize}
	\item {\tt \_XABORT\_RETRY}:
		When a {\tt xbegin} preemption point is created ({\tt save\_setjmp()}),
		if {\tt -S} is not set,
		both {\tt xabort\_codes} lists are initialized to contain this code.
		This represents the possibility for a hardware transaction to fail
		for reasons outside of the programmer's control such as system interrupts or cache eviction.
		If {\tt -S} is set they are initialized to empty,
		representing either STM's policy of failing only when a true memory conflict arises,
		or a TSX user wrapping all their {\tt \_xbegin()}s in a retry loop such as in \cite{sigbovik-htm}.
	\item {\tt \_XABORT\_CONFLICT}:
		Whenever DPOR detects a memory conflict between two preemption points ({\tt shimsham\_shm()}),
		if {\tt -A} is set,
		for each of the two which are within a transaction,
		it finds the nearest ancestor {\tt xbegin} preemption point by the same thread
		and adds this code to both lists.
	\item {\tt \_XABORT\_EXPLICIT}:
		Whenever the program invokes {\tt \_xabort()} ({\tt sched\_update\_user\_\allowbreak{}state\_machine()}),
		if {\tt -A} is set,
		this code is added to both lists, bitwise-ored with the user-supplied argument code
		as specified in \cite{htm-gcc}.
\end{itemize}

@@

@@

% ok so heres an example of why this is rly hard

>>>>

Up to three abort codes are considered depending on the testing options listed in the previous section.
All abort codes are simultaneously added to both lists,
unless already present on {\tt xabort\_codes\_ever},
in which case not added to {\tt xabort\_codes\_todo} to avoid duplicate work.
\begin{itemize}
	\item {\tt \_XABORT\_RETRY}:
		When a {\tt xbegin} preemption point is created ({\tt save\_setjmp()}),
		if {\tt -S} is not set,
		both {\tt xabort\_codes} lists are initialized to contain this code.
		This represents the possibility for a hardware transaction to fail
		for reasons outside of the programmer's control such as system interrupts or cache eviction.
		If {\tt -S} is set they are initialized to empty,
		representing either STM's policy of failing only when a true memory conflict arises,
		or a TSX user wrapping all their {\tt \_xbegin()}s in a retry loop such as in \cite{sigbovik-htm}.
	\item {\tt \_XABORT\_CONFLICT}:
		Whenever DPOR detects a memory conflict between two transitions ({\tt shimsham\_shm()}),
		if {\tt -A} is set,
		% for each of the two which are within a transaction, % no!!!!!!!
		%it finds the nearest ancestor {\tt xbegin} preemption point by the same thread
		if the later-executed transition is a transactional success path,
		it adds this code to both lists of the immediately preceding preemption point.
		Note that a transaction should suffer a conflict abort
		only when executed {\em after} a conflicting memory access
		to avoid circular causality
		(see {\tt 410user/progs/htm\_causality.c} for rationale).
		Should the transaction happen to be executed first,
		DPOR will first try to reorder it as normal,
		and then abort it as described if the conflict persists.
	\item {\tt \_XABORT\_EXPLICIT}:
		Whenever the program invokes {\tt \_xabort()} ({\tt sched\_update\_user\_\allowbreak{}state\_machine()}),
		if {\tt -A} is set,
		this code is added to both lists, bitwise-ored with the user-supplied argument code
		as specified in \cite{htm-gcc}.
\end{itemize}

@@

% don't forget to note that this only applies to xabort retry;

The proof of soundness for this optimization is left to future work;
while the basic idea is motivated by straightforward analogy to the known-sound sleep sets,
its intersection with Landslide's other exploration features may cause unforeseen problems.
Accordingly, I implemented the optimization as conservatively as possible,
falling back to no pruning in cases where how to prune soundly is not clear
(for example, when a second pair of conflicting transactions involving a third thread
interleaves while a first abort set is already being tracked).

@@

% TODO - rewrite the following example so that it only uses retry aborts
% ok so heres an example of why this is rly hard
